Preregistration Experiment 6: Is convergence more trustworthy when there are more choice options?

### 2023-06-05 Experiment_6/power_Exp6.Rmd
update experiment 6 before data collection

We set the parameters for convergence and the by-participant random effects based on experiment 4.

We further use the modeling results to produce the expected difference between the categories.

```{r}
# data from model
model_data <- read_csv("../experiment_4/data/model_3_options.csv") %>% 
  filter(options %in% c(3, 10)) %>% 
  mutate(
    # make numeric version of convergence
    convergence = case_when(constellation == "minority" ~ 0, 
                                 constellation == "dissensus" ~ 1, 
                                 constellation == "majority" ~ 2, 
                                 constellation == "consensus" ~ 3),
    # scale accuracy from 0 to 100 (currently from 0 to 1)
    average_accuracy = 100*average_accuracy,
    # scale competence reaching from 1 to 7 (currently from 0 to 1)
    average_relative_competence = (average_relative_competence * (7 - 1)) + 1,
    # make choice options a factor
    options = as.factor(options)
         )

# we run the model on averages from different iterations
model_accuracy <- lm(average_accuracy ~ convergence + options + convergence*options, data = model_data) %>% 
  tidy()

model_competence <- lm(average_relative_competence ~ convergence + options + convergence*options, data = model_data) %>% 
  tidy()
```

```{r}
# parameters for modelling accuracy 
beta_0   <- 20 # intercept
beta_c   <- 10 # effect of convergence when conflict of interest
beta_i   <- 5 # effect of independence (compared to conflict of interest)
beta_ci  <- 5 # difference in effect of convergence when number of options = 10 (compared to conflict of interest)
subj_0   <- 10 # by-subject random intercept sd
subj_c   <- 5 # by-subject random slope sd
subj_rho <- .2 # correlation between intercept and slope 
sigma    <- 5 # residual (error) sd (i.e. all varition that the model cannot account for)
```

We store a complete list of parameters in a list so that we can call them for functions later.

```{r}
# complete list of parameters (including some that will be introduced later)
parameters <- list(
  # outcome variable
  outcome = "accuracy", # takes names as input, e.g. "accuracy"
  # parameters for size of data frame
  n_subj = 10,              # n participants
  levels_convergence = 0:3,  # levels of convergence
  n_stimuli_per_level = 2, # n stimuli per participant
  # parameters for modelling outcome variable
  beta_0 = 40,   # intercept
  beta_c = 10,   # effect of convergence, 
  beta_i = 5,    # # effect of independence (compared to conflict of interest)
  beta_ci = 5, # difference in effect of convergence when independence (compared to conflict of interest)
  subj_0 = 10,   # by-subject random intercept sd
  subj_c = 5,   # by-subject random slope sd
  subj_rho = 0.5, # correlation between intercept and slope 
  sigma = 10     # residual (error) sd (i.e. all varition that the model cannot account for)
)
```

@@

# make data 
subjects <- data.frame(
  # add a subject id
  id = seq_len(n_subj),
  # add by participant pair of random effects
  subject_rfx, 
  # add `independence` condition
  independence = rep(c("conflict", "independent"), length.out = n_subj)
  ) %>% 
  # make a numeric variable to be able to generate competence
  mutate(independence_num = ifelse(independence == "conflict", 0, 1))
```

@@

#### Calculate accuracy values

@@

```{r}
dat_sim <- trials %>%
  mutate(accuracy = beta_0 + SU_0 + (beta_c + SU_c)*convergence + beta_i*independence_num + beta_ci*convergence*independence_num + e_ss,
         # truncate accuracy values so that they lie between 0 and 100 only
         accuracy = case_when(accuracy < 0 ~ 0, 
                              accuracy > 100 ~ 100,
                              TRUE ~ accuracy)) %>% 
  # retain only variables of interest
  select(id, convergence, independence, accuracy)
```

@@

```{r}
# independent condition
independent_plot <- ggplot(dat_sim %>% filter(independence == "independent"), 
       aes(x = as.factor(convergence), y = accuracy, 
           color = as.factor(convergence)
       )
) +
  # predicted means
  geom_hline(yintercept = beta_0) +
  geom_hline(yintercept = (beta_0 + beta_c + beta_ci)) +
  geom_hline(yintercept = (beta_0 + 2*(beta_c + beta_ci))) +
  geom_hline(yintercept = (beta_0 + 3*(beta_c + beta_ci))) +
  # actual data
  geom_violin(alpha = 0, show.legend = FALSE) +
  stat_summary(fun = mean, geom="crossbar", show.legend = FALSE) +
  scale_color_viridis_d() +
  ggtitle("Predicted versus simulated values") + 
  facet_wrap(~independence) +
  ylim(0, 100)

# conflict condition
conflict_plot <-ggplot(dat_sim %>% filter(independence == "conflict"), 
       aes(x = as.factor(convergence), y = accuracy, 
           color = as.factor(convergence)
           )
       ) +
  # predicted means
  geom_hline(yintercept = beta_0) +
  geom_hline(yintercept = (beta_0 + beta_c)) +
  geom_hline(yintercept = (beta_0 + 2*beta_c)) +
  geom_hline(yintercept = (beta_0 + 3*beta_c)) +
  # actual data
  geom_violin(alpha = 0, show.legend = FALSE) +
  stat_summary(fun = mean, geom="crossbar", show.legend = FALSE) +
  scale_color_viridis_d() +
  ggtitle("Predicted versus simulated values")  + 
  facet_wrap(~independence) +
  ylim(0, 100)

independent_plot + conflict_plot
```

@@

```{r}
# set up data generating function
draw_single_sample <- function(
  # outcome variable
  outcome, # takes names as input, e.g. "accuracy"
  # parameters for size of data frame
  n_subj,              # n participants
  levels_convergence,  # levels of convergence
  n_stimuli_per_level, # n stimuli per participant
  # parameters for modelling outcome variable
  beta_0,   # intercept
  beta_c,   # effect of convergence 
  beta_i = 5,    # # effect of independence (compared to conflict of interest)
  beta_ci = 5, # difference in effect of convergence when independence (compared to conflict of interest)
  subj_0,   # by-subject random intercept sd
  subj_c,   # by-subject random slope sd
  subj_rho, # correlation between intercept and slope 
  sigma     # residual (error) sd (i.e. all variation that the model cannot account for)
) {

@@

  # make data 
  subjects <- data.frame(
    # add a subject id
    id = seq_len(n_subj),
    # add by participant pair of random effects
    subject_rfx,
    # add `independence` condition
    independence = rep(c("conflict", "independent"), length.out = n_subj)
    ) %>% 
    # make a numeric variable to be able to generate competence
    mutate(independence_num = ifelse(independence == "conflict", 0, 1))

@@

  ## cross subject and item IDs
  sample <- expand_grid(subjects, stimuli)  %>%
    # add an error term
    # nrow(.) is the number of trials/observations, i.e. rows in the data frame
    mutate(e_ss = rnorm(nrow(.), mean = 0, sd = sigma)) %>%
    # simulate outcome variable
    mutate(outcome = beta_0 + SU_0 + (beta_c + SU_c)*convergence + 
             beta_i*independence_num + beta_ci*convergence*independence_num + e_ss, 
           outcome_variable = {{outcome}}) %>% 
    # retain only variables of interest
    select(id, convergence, independence, outcome, outcome_variable)

@@

```{r}
mod_sim <- lmer(accuracy ~ convergence + independence + 
                  independence*convergence + (1 + convergence | id), 
                dat_sim)
summary(mod_sim)
```

@@

```{r}
# get a tidy table of results
model_estimates <- broom.mixed::tidy(mod_sim) %>% 
  mutate_if(is.numeric, round, 3) %>%
  mutate(
    parameter = c("beta_0", "beta_c", "beta_i", "beta_ci", "subj_0", "subj_rho", "subj_c", "sigma"),
    value = c(beta_0, beta_c, beta_i, beta_ci, subj_0, subj_rho, subj_c, sigma),
  ) %>%
  select(effect:term, parameter, value, estimate:p.value)

@@

# estimates to extract
estimates = c("convergence", "independenceindependent", 
              "convergence:independenceindependent")

@@

  # estimate model for the drawn sample
  model_estimates <- lmer(outcome ~ convergence + independence + 
                            independence*convergence + (1 + convergence | id),
                          sample) %>% 
    # put into tidy output
    broom.mixed::tidy()

  # estimates to extract
  estimates = c("convergence", "independenceindependent", 
                "convergence:independenceindependent")

@@

```{r}
# list of parameters for accuracy
parameters_accuracy <- list(
  # outcome variable
  outcome = "accuracy", # takes names as input, e.g. "accuracy"
  # parameters for size of data frame
  sample_sizes  =  seq(from = 20, to = 200, by = 20), # different n's to simulate power for 
  levels_convergence = 0:3,  # levels of convergence
  n_stimuli_per_level = 2, # n stimuli per participant
  # parameters for modelling outcome variable
  beta_0 = 40,   # intercept
  beta_c = 5,   # effect of convergence 
  beta_i = 5,    # effect of independence (compared to conflict of interest)
  beta_ci = 5, # difference in effect of convergence when independence (compared to conflict of interest)
  subj_0 = 10,   # by-subject random intercept sd
  subj_c = 5,   # by-subject random slope sd
  subj_rho = -0.5, # correlation between intercept and slope 
  sigma = 10,     # residual (error) sd (i.e. all varition that the model cannot account for)
  # number of iterations (per sample size)
  iterations = 1000
)
```

@@

```{r}
# list of parameters for competence
parameters_competence <- list(
  # outcome variable
  outcome = "competence", # takes names as input, e.g. "accuracy"
  # parameters for size of data frame
  sample_sizes  =  seq(from = 20, to = 300, by = 20), # different n's to simulate power for 
  levels_convergence = 0:3,  # levels of convergence
  n_stimuli_per_level = 2, # n stimuli per participant
  # parameters for modelling outcome variable
  beta_0 = 3.5,   # intercept
  beta_c = 0.25,   # effect of convergence 
  beta_i = 0.25,    # # effect of independence (compared to conflict of interest)
  beta_ci = 0.5, # difference in effect of convergence when independence (compared to conflict of interest)
  subj_0 = 1,   # by-subject random intercept sd
  subj_c = 0.25,   # by-subject random slope sd
  subj_rho = -0.5, # correlation between intercept and slope 
  sigma = 1,     # residual (error) sd (i.e. all varition that the model cannot account for)
  # number of iterations (per sample size)
  iterations = 1000
)
```

@@

>>>>

We set the parameters for the by-participant random effects and the error term based on the estimates from experiment 4.

```{r, message=FALSE}
# read data
d <- read_csv("../experiment_4/data/cleaned.csv")

# get random effects for accuracy

# calculate model
model_accuracy <- lmer(accuracy ~ convergence + (1 + convergence | id), 
                                 data = d)

# store estimates with according parameters in list
random_effects_accuracy <- tidy(model_accuracy, effects = "ran_pars") %>% 
  mutate(parameter = case_when(term == "sd__(Intercept)" ~ "subj_0", 
                               term == "cor__(Intercept).convergence" ~ "subj_rho",
                               term == "sd__convergence" ~ "subj_c", 
                               term == "sd__Observation" ~ "sigma")
         ) %>% 
  select(parameter, estimate)

# get random effects for competence

# calculate model
model_competence <- lmer(competence ~ convergence + 
                           (1 + convergence | id), data = d)

# store estimates with according parameters in list
random_effects_competence <- tidy(model_competence, effects = "ran_pars") %>% 
  mutate(parameter = case_when(term == "sd__(Intercept)" ~ "subj_0", 
                               term == "cor__(Intercept).convergence" ~ "subj_rho",
                               term == "sd__convergence" ~ "subj_c", 
                               term == "sd__Observation" ~ "sigma")
         ) %>% 
  select(parameter, estimate)

We further use the modeling results to set the parameters for all fixed effects.

```{r, message=FALSE}
# data from model
model_data <- read_csv("../experiment_4/data/model_vary_options_3_sample.csv") %>% 
  filter(options %in% c(3, 10)) %>% 
  mutate(
    # make numeric version of convergence
    convergence = case_when(constellation == "minority" ~ 0, 
                                 constellation == "dissensus" ~ 1, 
                                 constellation == "majority" ~ 2, 
                                 constellation == "consensus" ~ 3),
    # scale accuracy from 0 to 100 (currently from 0 to 1)
    average_accuracy = 100*average_accuracy,
    # scale competence reaching from 1 to 7 (currently from 0 to 1)
    average_relative_competence = (average_relative_competence * (7 - 1)) + 1,
    # make choice options a factor
    options = as.factor(options)
         )

# we run the model on averages from different iterations

# for accuracy
model_accuracy <- lm(average_accuracy ~ convergence + options + convergence*options, data = model_data)

# store estimates with according parameters in list
fixed_effects_accuracy <- tidy(model_accuracy) %>% 
  mutate(parameter = case_when(term == "(Intercept)" ~ "beta_0", 
                               term == "convergence" ~ "beta_c",
                               term == "options10" ~ "beta_n",
                               term == "convergence:options10" ~ "beta_cn")
         ) %>% 
  select(parameter, estimate)

# for competence
model_competence <- lm(average_relative_competence ~ convergence + options + convergence*options, data = model_data)

# store estimates with according parameters in list
fixed_effects_competence <- tidy(model_competence) %>% 
  mutate(parameter = case_when(term == "(Intercept)" ~ "beta_0", 
                               term == "convergence" ~ "beta_c",
                               term == "options10" ~ "beta_n",
                               term == "convergence:options10" ~ "beta_cn")
         ) %>% 
  select(parameter, estimate)
```

```{r}
# parameters for modelling accuracy 
parameters_accuracy <- bind_rows(random_effects_accuracy, fixed_effects_accuracy)
# split() creates a list using the values in a dataframe column
parameters_accuracy <- split(parameters_accuracy$estimate, parameters_accuracy$parameter)

# parameters for modelling competence 
parameters_competence <- bind_rows(random_effects_competence, fixed_effects_competence)
parameters_competence <- split(parameters_competence$estimate, parameters_competence$parameter)

parameters_accuracy
parameters_competence
```

To these effect parameters, we have to add a set of additional parameters. We store a complete list of parameters in a list so that we can call them for functions later. Note that here we'll only set parameters for accuracy, for demonstration purposes. We'll do the same for competence later when running the actual simulations.

```{r}
# store parameters in separate objects to be able to call them for demonstration purposes
# (for accuracy only)
beta_0   <- parameters_accuracy$beta_0 # intercept
beta_c   <- parameters_accuracy$beta_c # effect of convergence when 3 choice options
beta_n   <- parameters_accuracy$beta_n # effect of 10 options (vs. 3) when convergence = 0
beta_cn  <- parameters_accuracy$beta_cn # difference in effect of convergence when 10 options (vs. 3)
subj_0   <- parameters_accuracy$subj_0 # by-subject random intercept sd
subj_c   <- parameters_accuracy$subj_c # by-subject random slope sd
subj_rho <- parameters_accuracy$subj_rho # correlation between intercept and slope 
sigma    <- parameters_accuracy$sigma # residual (error) sd (i.e. all varition that the model cannot account for)

# complete list of parameters (including some that will be introduced later)
parameters <- c(
  parameters_accuracy,
  list(
  # outcome variable
  outcome = "accuracy", # takes names as input, e.g. "accuracy"
  # parameters for size of data frame
  n_subj = 100,              # n participants
  levels_convergence = 0:3,  # levels of convergence
  n_stimuli_per_level = 2 # n stimuli per participant
  )
  )
```

@@

# make data 
subjects <- data.frame(
  # add a subject id
  id = seq_len(n_subj),
  # add by participant pair of random effects
  subject_rfx, 
  # add `independence` condition
  number_options = rep(c("3", "10"), length.out = n_subj)
) %>% 
  # make a numeric variable to be able to generate competence
  mutate(number_options_num = ifelse(number_options == "3", 0, 1), 
         # relevel the number_options variable
         number_options = relevel(as.factor(number_options), ref = "3"))
```

@@

### Calculate accuracy values

@@

```{r}
dat_sim <- trials %>%
  mutate(accuracy = beta_0 + SU_0 + (beta_c + SU_c)*convergence + beta_n*number_options_num + beta_cn*convergence*number_options_num + e_ss,
         # truncate accuracy values so that they lie between 0 and 100 only
         accuracy = case_when(accuracy < 0 ~ 0, 
                              accuracy > 100 ~ 100,
                              TRUE ~ accuracy)) %>% 
  # retain only variables of interest
  select(id, convergence, number_options, accuracy)
```

@@

```{r}
# independent condition
options_10_plot <- ggplot(dat_sim %>% filter(number_options == "10"), 
       aes(x = as.factor(convergence), y = accuracy, 
           color = as.factor(convergence)
       )
) +
  # predicted means
  geom_hline(yintercept = beta_0) +
  geom_hline(yintercept = (beta_0 + beta_c + beta_cn)) +
  geom_hline(yintercept = (beta_0 + 2*(beta_c + beta_cn))) +
  geom_hline(yintercept = (beta_0 + 3*(beta_c + beta_cn))) +
  # actual data
  geom_violin(alpha = 0, show.legend = FALSE) +
  stat_summary(fun = mean, geom="crossbar", show.legend = FALSE) +
  scale_color_viridis_d() +
  ggtitle("Predicted versus simulated values") + 
  facet_wrap(~number_options) +
  ylim(0, 120)

# conflict condition
options_3_plot <-ggplot(dat_sim %>% filter(number_options == "3"), 
       aes(x = as.factor(convergence), y = accuracy, 
           color = as.factor(convergence)
           )
       ) +
  # predicted means
  geom_hline(yintercept = beta_0) +
  geom_hline(yintercept = (beta_0 + beta_c)) +
  geom_hline(yintercept = (beta_0 + 2*beta_c)) +
  geom_hline(yintercept = (beta_0 + 3*beta_c)) +
  # actual data
  geom_violin(alpha = 0, show.legend = FALSE) +
  stat_summary(fun = mean, geom="crossbar", show.legend = FALSE) +
  scale_color_viridis_d() +
  ggtitle("Predicted versus simulated values")  + 
  facet_wrap(~number_options) +
  ylim(0, 120)

options_10_plot + options_3_plot
```

@@

```{r}
# set up data generating function
draw_single_sample <- function(
  # outcome variable
  outcome, # takes names as input, e.g. "accuracy"
  # parameters for size of data frame
  n_subj,              # n participants
  levels_convergence,  # levels of convergence
  n_stimuli_per_level, # n stimuli per participant
  # parameters for modelling outcome variable
  beta_0,   # intercept
  beta_c,   # effect of convergence 
  beta_n,  # effect of independence (compared to conflict of interest)
  beta_cn,  # difference in effect of convergence when independence (compared to conflict of interest)
  subj_0,   # by-subject random intercept sd
  subj_c,   # by-subject random slope sd
  subj_rho, # correlation between intercept and slope 
  sigma     # residual (error) sd (i.e. all variation that the model cannot account for)
) {

@@

  # make data 
  subjects <- data.frame(
    # add a subject id
    id = seq_len(n_subj),
    # add by participant pair of random effects
    subject_rfx,
    # add `independence` condition
    number_options = rep(c("3", "10"), length.out = n_subj)
    ) %>% 
    # make a numeric variable to be able to generate competence
    mutate(number_options_num = ifelse(number_options == "3", 0, 1), 
           # relevel the number_options variable
           number_options = relevel(as.factor(number_options), ref = "3")
           )

@@

  ## cross subject and item IDs
  sample <- expand_grid(subjects, stimuli)  %>%
    # add an error term
    # nrow(.) is the number of trials/observations, i.e. rows in the data frame
    mutate(e_ss = rnorm(nrow(.), mean = 0, sd = sigma)) %>%
    # simulate outcome variable
    mutate(outcome = beta_0 + SU_0 + (beta_c + SU_c)*convergence + 
             beta_n*number_options_num + beta_cn*convergence*number_options_num + e_ss, 
           outcome_variable = {{outcome}}) %>% 
    # retain only variables of interest
    select(id, convergence, number_options, outcome, outcome_variable)

@@

```{r}
mod_sim <- lmer(accuracy ~ convergence + number_options + 
                  number_options*convergence + (1 + convergence | id), 
                dat_sim)
summary(mod_sim)
```

@@

```{r}
# get a tidy table of results
model_estimates <- broom.mixed::tidy(mod_sim) %>% 
  mutate_if(is.numeric, round, 3) %>%
  mutate(
    parameter = c("beta_0", "beta_c", "beta_n", "beta_cn", "subj_0", "subj_rho", "subj_c", "sigma"),
    value = c(beta_0, beta_c, beta_n, beta_cn, subj_0, subj_rho, subj_c, sigma),
  ) %>%
  select(effect:term, parameter, value, estimate:p.value)

@@

# estimates to extract
estimates = c("convergence", "number_options10", 
              "convergence:number_options10")

@@

  # estimate model for the drawn sample
  model_estimates <- lmer(outcome ~ convergence + number_options + 
                            number_options*convergence + (1 + convergence | id),
                          sample) %>% 
    # put into tidy output
    broom.mixed::tidy()

  # estimates to extract
  estimates = c("convergence", "number_options10", 
              "convergence:number_options10")

@@

```{r}
# list of parameters for accuracy
parameters_accuracy <- c(parameters_accuracy, 
                         list(
                           # outcome variable
                           outcome = "accuracy", # takes names as input, e.g. "accuracy"
                           # parameters for size of data frame
                           sample_sizes  =  seq(from = 20, to = 200, by = 20), # different n's to simulate power for 
                           levels_convergence = 0:3,  # levels of convergence
                           n_stimuli_per_level = 2, # n stimuli per participant
                           # number of iterations (per sample size)
                           iterations = 1000
                           )
                         )
```

@@

```{r}
# list of parameters for competence
parameters_competence <- c(parameters_competence, 
                         list(
                           # outcome variable
                           outcome = "competence", # takes names as input, e.g. "competence"
                           # parameters for size of data frame
                           sample_sizes  =  seq(from = 20, to = 300, by = 20), # different n's to simulate power for 
                           levels_convergence = 0:3,  # levels of convergence
                           n_stimuli_per_level = 2, # n stimuli per participant
                           # number of iterations (per sample size)
                           iterations = 1000
                           )
                         )
```

@@

### Calculate precise sample size threshold

```{r}
# accuracy
power_accuracy %>% filter(term == "convergence:number_options10" & 
                            power >= 0.9) %>%
  arrange(power) %>% 
  slice(1)
```

```{r}
# competence
power_competence %>% filter(term == "convergence:number_options10") %>%
  arrange(desc(power)) 
```

### 2023-06-05 Experiment_6/preregistration.Rmd
update experiment 6 before data collection

---
title: 'Preregistration Experiment 6: Is convergence more trustworthy when there are more choice options?'
author: "Jan Pfänder, Hugo Mercier"
date: "2023-03-08"
output:
  html_document:
    keep_md: yes
  pdf_document: 
      keep_md: yes
---

@@

# simulate data
data <- tibble(
  convergence = sample(0:3, replace = TRUE, size=100),
  accuracy = sample(0:100, replace = TRUE, size=100),
  competence = sample(1:7, replace = TRUE, size=100),
  options = rep(c("3", "10"), each = 50),
  id = rep(1:5, length.out = 100)
  ) 
```

@@

**Options**. Number of options has two levels: (i) '3' and (ii) '10'.

We manipulate independence between participants, i.e. each participant gets assigned to either the '3' or the '10' condition. All stimuli that participants see in their respective condition will involve the same number of choice options.

@@

Considering only `3` options condition of the experiment at hand, we expect the findings of experiment four to replicate. In experiment four, players would choose among `3` options.

@@

### H1b:  In the three options condition, participants perceive an independent informant as more competent the more their estimate converges with the estimates of other informants, given three choice options.

@@

How about a context in which informants (i.e. the fictive players) could choose not among 3, but among 10 options? In line with our findings from experiment three, we predict that effects of convergence are more positive when there are more choice options:

@@

@@

## Alternative stimuli with greater distance

We kept the same distance between 10 and 3. That's conservative - people might think that other options are not relevant and hence don't make a difference between 3 and 10 options. We therefore run the hypotheses set two comparing 3 to a version of 10 with larger distance.

@@

We ran a power simulation to inform our choice of sample size. All assumptions and details on the procedure can be found in the `power_Exp5.Rmd` document. We ran two different power analyses, one for each outcome variable. We set the power threshold for our experiment to 90%.

The power simulation for `accuracy` suggested that for 80 participants, we would have a power of at least 90% for the interaction effect. The simulation for `competence` suggested that with already 40 participants, we would detect an interaction, but only with 60 participants we also detect an effect of convergence.

However, due to uncertainty about our assumptions and because we are anticipating failed attention checks, *we will recruit a sample of `200` participants*.

```{r accuracy-image, echo=FALSE, fig.cap="Results of power simulation for competence"}
#knitr::include_graphics("power_Exp5_files/figure-html/plot-power-accuracy-1.png")
```

```{r competence-image, echo=FALSE, fig.cap="Results of power simulation for competence"}
#knitr::include_graphics("power_Exp5_files/figure-html/plot-power-competence-1.png")
```

>>>>

---
title: 'Preregistration Experiment 6: Is convergence more trustworthy when there are more choice options?'
author: "Jan Pfänder, Hugo Mercier"
date: "2023-03-08"
output:
  pdf_document: 
      keep_md: yes
  html_document:
    keep_md: yes
---

@@

# simulate data
data <- tibble(
  convergence = sample(0:3, replace = TRUE, size=100),
  accuracy = sample(0:100, replace = TRUE, size=100),
  competence = sample(1:7, replace = TRUE, size=100),
  condition = sample(c("3", "10", "10_distant"), replace = TRUE, 
                   prob = c(0.5, 0.25, 0.25), size = 100),
  id = rep(1:5, length.out = 100)
  ) %>% 
  mutate(stimuli_10_version = ifelse(str_detect(condition, "3"), NA, 
                                     condition), 
         options = ifelse(str_detect(condition, "10"), "10", 
                          condition)
                                     )
```

@@

**Number of choice options**. Number of options has two levels: '3' '10'.

We manipulate number of choice options between participants, i.e. each participant gets assigned to either the '3' or the '10' condition. All stimuli that participants see in their respective condition will involve the same number of choice options.

As for the '10' options condition, participants will see one of two distinct set of stimuli (`stimuli_10_version`): one in which the range of the answers corresponds to the range of the '3' options condition (`stimuli_10_version == 10`), and another with increased range (`stimuli_10_version == 10_distant`; see Appendix). We add the increased range condition because we think that there is the possibility that participant might not consider all options as relevant when they only see scenarios in which all answers cluster. For the main analyses, we will simply merge these two stimuli sets and treat them as part of the same '10' options condition. However, we will investigate potential differences as a research question.

@@

Considering only `3` options condition of the experiment at hand, we expect the findings of experiment four to replicate, for it is the exact same experiment as experiment four in this case.

@@

### H1b:  In the three options condition, participants perceive an independent informant as more competent the more their estimate converges with the estimates of other informants.

@@

How about a context in which informants (i.e. the fictive players) could choose not among 3, but among 10 options? Following the results from our model, we predict that effects of convergence are more positive when there are more choice options:

@@

## Research question

### RQ1: Within the 10 choice options condition, is the effect of convergence more positive for the set of stimuli with greater distance?

# random intercept and slope by participants
model_accuracy <- lmer(accuracy ~ convergence + stimuli_10_version + 
                            stimuli_10_version*convergence + (1 + convergence | id), 
                       data = data)

# in case of non-convergence: random intercept by participants only
alt_model_accuracy <- lmer(accuracy ~ convergence + stimuli_10_version + 
                            stimuli_10_version*convergence + (1 | id), 
                           data = data)

# models for competence

# random intercept and slope by participants
model_competence <- lmer(competence ~ convergence + stimuli_10_version + 
                            stimuli_10_version*convergence + (1 + convergence | id), 
                       data = data)

# in case of non-convergence: random intercept by participants only
alt_model_competence <- lmer(competence ~ convergence + stimuli_10_version + 
                            stimuli_10_version*convergence + (1 | id), 
                           data = data)
```

Should we find the difference, we will re-run the analyses of the second set of hypotheses restraining the data to the subset of the more distant set of stimuli for the 10 options condition.

@@

@@

We ran a power simulation to inform our choice of sample size. All assumptions and details on the procedure can be found in the `power_Exp6.Rmd` document. We used previous experiments and estimates of our models to inform our choice of parameter values.

We ran two different power analyses, one for each outcome variable. We set the power threshold for our experiment to 90%.

The power simulation for `accuracy` suggested that for 140 participants we would cross the power threshold of 90% for the interaction effect (power = 0.928). The simulation for `competence` suggested that with 300 participants, we would detect an interaction with a power of 87% (power = 0.872).

Due to budget constraints, we will consider a sample of 300 participants as good enough in terms of power.

```{r accuracy-image, echo=FALSE, fig.cap="Results of power simulation for competence"}
knitr::include_graphics("power_Exp6_files/figure-html/plot-power-accuracy-1.png")
```

```{r competence-image, echo=FALSE, fig.cap="Results of power simulation for competence"}
knitr::include_graphics("power_Exp6_files/figure-html/plot-power-competence-1.png")
```
